---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---


## Loading and preprocessing the data

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(lattice)
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
data1 <- data %>% group_by(date) %>% summarise(steps=sum(steps))
hist(data1$steps, main="Histogram of daily steps", xlab="")
```
### The mean and median of the total number of steps taken per day is:
```{r}
summary(data1$steps)
```

## What is the average daily activity pattern?

```{r}
data2 <- data %>% group_by(interval) %>% summarise(steps=mean(steps, na.rm=TRUE))
data2$time <- lapply(data2$interval, function(x){return (paste(as.character(x%/%100), as.character(x%%100), sep=":"))})
data2$time <- strptime(data2$time, format="%H:%M")

plot(data2$time, data2$steps, type="n", xlab="", ylab="Steps", main="Average daily activity pattern")
lines(data2$time, data2$steps)
```

### The interval that contains the maximum number of steps, on average, is:

```{r}
print((data2 %>% filter(data2$steps == max(data2$steps)))$interval)
```


## Imputing missing values
### The total number of missing values in the dataset is:
```{r}
summary(is.na(data$steps))
```

### I created a new dataset filling the NAs with the mean for the 5 minute interval

```{r}
data3<-data
data3$steps <- invisible(sapply(seq_along(data$steps), function(x){
  if(is.na(data[x,1])){
    return (pull(data2[data[x, 3]==data2$interval, 2]))
  }else{
    return (data[x,1])
  }
}))
```

### These are the histogram and the mean and median total number of steps taken per day. The values are similar.

```{r}
data32 <- data3 %>% group_by(date) %>% summarise(steps=sum(steps))
hist(data32$steps, main="Histogram of daily steps", xlab="")
summary(data32$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r, warning=FALSE, message=FALSE}
data4 <- data3
data4$time <- sapply(data4$interval, function(x){return (paste(as.character(x%/%100), as.character(x%%100), sep=":"))})
data4$fulltime <- paste(data4$date, data4$time)
data4$fulldate <- strptime(data4$fulltime, format="%Y-%m-%d %H:%M")
data4$dayOfTheWeek <- weekdays(data4$fulldate)
data4$dayOfTheWeek <- sapply(data4$dayOfTheWeek, function(x){return (ifelse(x=="sábado" | x=="domingo", "weekend", "weekday"))})
data4$dayOfTheWeek <- factor(x=data4$dayOfTheWeek, levels=c("weekday", "weekend"), labels=c("weekday", "weekend"))
data4 <- data4 %>% group_by(dayOfTheWeek, interval) %>% summarise(steps=sum(steps)) 

xyplot(steps ~ interval | dayOfTheWeek, data=data4, type="l")

```